\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{float}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	urlcolor=blue,
}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{color}
\usepackage{setspace}
\usepackage{subfigure}
\usepackage{algorithm, algorithmic}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{url}
\usepackage{cite}
\usepackage{fancyvrb}
\usepackage{amsmath, amsfonts}
\usepackage{rotating}
\input{definitions}

\begin{document}
\sloppy

\title{TBD}

\author{\IEEEauthorblockN{David B.~Knoester and others}
\IEEEauthorblockA{Dept.~of Microbiology and Molecular Genetics\\
Michigan State University\\
East Lansing, Michigan 48824\\
Email: \{dk\}@msu.edu}}
\maketitle

\begin{abstract}
Combining systems that self-organize with those that self-adapt is an important problem in SASO systems.  While cellular automata have long been used as a system to study self-organization, common rule-based approaches describing the functionality of a cellular automata do not provide a capability for self-adaptation.
%
In this paper, we describe an approach to cellular automata where the rule table is replaced by state machines that have been evolved via a novel evolutionary algorithm.  Through a number of different experiments, we show how this evolutionary algorithm is able to discover state machines that solve the well-known majority bit problem in 1-, 2-, and 3-dimensional cellular automata.
%
We then show that probabilistic state machine are capable of self-adaptation, and that adaptive PSMs that have been evolved for the majority bit problem more readily adapt to new problems, i.e., minority bit.
%
Finally, we show that when the state machines controlling each cell in a cellular automata are provided with information regarding the global state of the automata, performance, evolvability, and scalability of the resulting solutions dramatically improve.
%
Taken together, these results strengthen the case for cellular automata as an ideal study system for self-organization, indicate CAs for further exploratory work on SASO systems, and suggest that some form of reinforcement might supersede self-adaptation as a key component of SASO systems.
\end{abstract}

\begin{IEEEkeywords}
Evolutionary algorithm, cellular automata, self-organization, self-adaptation, reinforcement learning.
\end{IEEEkeywords}

\input{s1-introduction}
\input{s2-relatedWork}
\input{s3-methods}
\input{s4-results}

\section{Conclusion}\label{s:conclusion}
We have described one way in which self-organizing and self-adaptive systems can be combined to solve problems.  While the experiments described in this paper use a model system (a cellular automata), the techniques employed are general.  In addition to presenting results that highlight the ability of EAs to discover SASO systems, we have also shown the importance of a globally-defined RL signal.  In future work, we plan to explore different topologies of CA, including those in which the automata themselves are able to adjust the underlying connectivity pattern of the CA.  We also intend to explore in more detail the relationship between an externally-defined RL signal and the capabilities of an adaptive agent.

\section*{Further Information and Acknowledgments}
Source code and data for the experiments described in this paper is freely available, and is supplied as-is.  Source code: \url{?}, data: \url{?}.


%\bibliography{dk}
\bibliographystyle{IEEEtran}
\end{document}
